---
title: "President of the LDS Church Succession"
output:
  html_document:
    df_print: paged
---

This analysis addresses succession in the presidency of the LDS Church. See Github for <a href="https://github.com/nwstephens/lds" target="_blank">my code</a>. See <a href="http://zelophehadsdaughters.com/2018/01/04/church-president-probability-changes-with-president-monsons-death/#more-12559" target="_blank">here</a> for a similiar analysis.

## Setup

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(tibble)
library(purrr)
library(rvest)
```

## Data

#### Current members

Current members of the <a href="https://en.wikipedia.org/wiki/List_of_general_authorities_of_The_Church_of_Jesus_Christ_of_Latter-day_Saints#Quorum_of_the_Twelve_Apostles" target="_blank">Quorum of the Twelve</a>.

```{r}
rawdat <- matrix(
  nrow = 15, 
  byrow = TRUE, 
  dimnames = list(1:15, c("rnk", "name", "dob", "dos")), 
  data=c(
   "1",    "Russell M. Nelson",          "1924-09-09",    "1984-04-12",
   "2",    "Dallin H. Oaks",             "1932-08-12",    "1984-05-03",
   "3",    "Jeffrey R. Holland",         "1940-12-03",    "1994-06-23",
   "4",    "Henry B. Eyring",            "1933-05-31",    "1995-04-06",
   "5",    "Dieter F. Uchtdorf",         "1940-11-06",    "2004-10-07",
   "6",    "David A. Bednar",            "1952-06-15",    "2004-10-07",
   "7",    "Quentin L. Cook",            "1940-09-08",    "2007-10-11",
   "8",    "D. Todd Christofferson",     "1945-01-24",    "2008-04-10",
   "9",    "Neil L. Andersen",           "1951-08-09",    "2009-04-09",
  "10",    "Ronald A. Rasband",          "1951-02-06",    "2015-10-08",
  "11",    "Gary E. Stevenson",          "1955-08-06",    "2015-10-08",
  "12",    "Dale G. Renlund",            "1952-11-13",    "2015-10-08",
  "13",    "Gerrit W. Gong",             "1953-12-23",    "2018-03-31",
  "14",    "Ulisses Soares",             "1958-10-02",    "2018-03-31",
  "15",    "Patrick Kearon",             "1961-07-18",    "2023-12-07"  # new update since last version (2021)
 )
)
```

#### Organize data

Format dates and create current age based on today's date.

```{r}
today <- as.Date(format(Sys.Date(), format="%Y-%m-%d"))
dat <- as_tibble(rawdat) %>%
  mutate(dob = as.Date(dob)) %>%
  mutate(dos = as.Date(dos)) %>%
  mutate(age = today - dob)
```

#### Import life table

There are many life tables to chose from, and preliminary analysis shows that the predicted probabilities are highly sensitive depending on which table is chosen. This table is taken from the Social Security administration.

```{r}
lifetable <- read_html("https://www.ssa.gov/oact/STATS/table4c6.html") %>%
  html_nodes(xpath="//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]") %>%
  html_text %>%
  as.numeric %>%
  data.frame(age = 0:119, pdeath = .)
```

## Predictions

Solve the question: What is the probability of becoming president for each member of the quorum based solely on average time until death.

#### Inputs

```{r}
# Simulation inputs
N <- 20000
years <- 45

# Data inputs
age <- floor(dat$age/365)
nams <- dat$name
pvec <- lifetable$pdeath
```

#### Simulation

```{r}
# Probabilities over time
pmat <- outer(age, 1:years, function(x, y) pvec[x + y])
pmat[is.na(pmat)] <- 0
pmat[, 1] <- 0 # assumes no one dies in the first year

# Simulate
out <- matrix("", years, N)
for(i in 1:N){
  alive <- apply(pmat, 1, function(x) !cumsum(rbinom(years, 1, x)))
  pres <- apply(cbind(alive, TRUE), 1, which.max)
  out[, i] <- c(nams, "none")[pres]
}
```

#### Average predicted probabilities

```{r}
pct <- map(nams, function(x) apply(out == x, 2, any)) %>%
  map_dbl(mean) * 100
```

#### Average tenure

```{r}
tenure <- unlist(apply(out, 2, table))
avg_tenure <- unlist(tapply(tenure, names(tenure), mean))
```

## Summary

### Table

```{r}
dat2 <- dat %>%
  mutate(pct = pct) %>%
  mutate(tenure = round(as.numeric(avg_tenure[nams]), 2))
dat2
```

#### Visualize

```{r}
ggplot(dat2, aes(reorder(nams, 15:1), pct)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  theme_minimal() +
  coord_flip() +
  labs(x = "", y = "Percent", title = "Likelihood of becoming president")
ggsave("pred_prob.png")
```

