---
title: "President of the LDS Church Succession"
output: html_notebook
---

This analysis addresses succession in the presidency of the LDS Church. See Github for [code](https://github.com/nwstephens/lds.git). For a comporable analysis see [Church President Probability Changes with President Monson’s Death](http://zelophehadsdaughters.com/2018/01/04/church-president-probability-changes-with-president-monsons-death/#more-12559).

## Setup

```{r, message=FALSE}
library(tidyverse)
library(rvest)
```

## Current members

Current members of the [Quorum of the Twelve](https://en.wikipedia.org/wiki/List_of_general_authorities_of_The_Church_of_Jesus_Christ_of_Latter-day_Saints#Quorum_of_the_Twelve_Apostles).

```{r}
rawdat <- matrix(
  nrow = 13, 
  byrow = TRUE, 
  dimnames = list(1:13, c("rnk", "name", "dob", "dos")), 
  data=c(
   "1",    "Russell M. Nelson",          "1924-09-09",    "1984-04-12",
   "2",    "Dallin H. Oaks",             "1932-08-12",    "1984-05-03",
   "3",    "M. Russell Ballard",         "1928-10-08",    "1985-10-10",
   "4",    "Jeffrey R. Holland",         "1940-12-03",    "1994-06-23",
   "5",    "Henry B. Eyring",            "1933-05-31",    "1995-04-06",
   "6",    "Dieter F. Uchtdorf",         "1940-11-06",    "2004-10-07",
   "7",    "David A. Bednar",            "1952-06-15",    "2004-10-07",
   "8",    "Quentin L. Cook",            "1940-09-08",    "2007-10-11",
   "9",    "D. Todd Christofferson",     "1945-01-24",    "2008-04-10",
  "10",    "Neil L. Andersen",           "1951-08-09",    "2009-04-09",
  "11",    "Ronald A. Rasband",          "1951-02-06",    "2015-10-08",
  "12",    "Gary E. Stevenson",          "1955-08-06",    "2015-10-08",
  "13",    "Dale G. Renlund",            "1952-11-13",    "2015-10-08"
  )
)
```

## Organize data

Format dates and create current age based on today's date.

```{r}
today <- as.Date("2018-01-08")
dat <- as_tibble(rawdat) %>%
  mutate(dob = as.Date(dob)) %>%
  mutate(dos = as.Date(dos)) %>%
  mutate(age = today - dob)
dat
```

## Import life table

There are many life tables to chose from, and preliminary analysis shows that the predicted probabilities are highly sensitive depending on which table is chosen. This table is taken from the Social Security administration.

```{r}
lifetable <- read_html("https://www.ssa.gov/oact/STATS/table4c6.html") %>%
  html_nodes(xpath="//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]") %>%
  html_text %>%
  as.numeric %>%
  data.frame(age = 0:119, pdeath = .)
```

## Simulation

Solve the question: What is the probability of becoming president for each member of the quorum based solely on average time until death.

```{r}
N <- 10000
years <- 35

age <- floor(dat$age/365)
nams <- dat$name
pvec <- lifetable$pdeath

pmat <- outer(age, 1:years, function(x, y) pvec[x + y])
pmat[is.na(pmat)] <- 0
pmat[, 1] <- 0

out <- matrix("", years, N)
for(i in 1:N){
  alive <- apply(pmat, 1, function(x) !cumsum(rbinom(years, 1, x)))
  pres <- apply(cbind(alive, TRUE), 1, which.max)
  out[, i] <- c(nams, "none")[pres]
}

pct <- map(nams, function(x) apply(out == x, 2, any)) %>%
  map_dbl(mean) * 100
```

## Plot

Visualize the predicted probabilities.

```{r}
dat %>%
  mutate(pct = pct) %>%
  ggplot(aes(reorder(nams, 13:1), pct)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  theme_minimal() +
  coord_flip() +
  labs(x = "", y = "Percent", title = "Likelihood of becoming president")
```


